---
title: "Univariate Testing"
author: "Eric Kramer"
output: html_document
runtime: shiny
---

```{r echo=F, include=FALSE}
library("data.table")
library("dplyr")
library("ggvis")
library("shiny")


load("./data/univariate_performance.Rdata")
load("./data/census_training.Rdata")
load("./data/missingness.Rdata")

census_sample = census_training %>%
  as.data.frame %>%
  group_by(Target) %>%
  sample_n(5000)  # downsampling for visualization

performance = missingness %>%
  select(Variable, Missing_Frequency, Rare_Frequency) %>%
  as.data.frame %>%
  inner_join(performance)

```

## Note on Preprocessing


I  downsampled the dataset for visualization, taking 1,000 samples from each income category. This improves visual comparison between income groups. Otherwise, the counts are too imbalanced to make easy judgements.

For categorical variables, I set all values with a frequency less than 5% to *rare*. This mostly affected the variable capturing county-of-birth information and detailed household / employment statistics statistics. We could attempt to recover some of the information in these variables by mapping countries to regions and reducing the number of household categories.


## AUC measurements for each variable

I tested the predictive capability of each variable using a random forest. A separate random forest was grown for each variable, and its performance was measured using an out-of-bounds AUC measurement. 



```{r echo=F}
performance %>%
  mutate(AUC=as.numeric(AUC)) %>%
  select(Variable, AUC, Missing_Frequency, Rare_Frequency) %>%
  arrange(-AUC) %>%
  renderDataTable(options=list(pageLength = 10))
```

## Insights from Univariate Analysis

### MajorOccupation: Executives and Professionals earn more
```{r echo=F}
census_sample %>%
  ggvis(~MajorOccupation, fill=~Target)  %>%
  layer_bars() %>%
  add_axis("x",
           title="",
         title_offset=100,
         properties=axis_props(labels=list(angle=45,
                                           align="left")))
```

### Education: Individuals with higher education earn more
```{r echo=F}
census_sample %>%
  ggvis(~Education, fill=~Target)  %>%
  layer_bars() %>%
  add_axis("x",
           title="",
         title_offset=100,
         properties=axis_props(labels=list(angle=45,
                                           align="left")))
```

### MaritalStat: Married households earn more
```{r echo=F}
census_sample %>%
  ggvis(~MaritalStat, fill=~Target)  %>%
  layer_bars() %>%
  add_axis("x",
           title="",
         title_offset=100,
         properties=axis_props(labels=list(angle=45,
                                           align="left")))
```


### Sex: We also see evidence for the *wage-gap* -- men earn more than women
```{r echo=FALSE}
census_sample %>%
  ggvis(~Sex, fill=~Target)  %>%
  layer_bars() %>%
  add_axis("x",
           title="",
         title_offset=100,
         properties=axis_props(labels=list(angle=45,
                                           align="left")))   
```

### Age: Workers in prime working-age (30 to 60 years old) earn the most 
```{r echo=F, fig.height=7}
census_sample %>% 
  ggvis(~Age, fill=~Target) %>%
  group_by(Target) %>%
  layer_histograms(width=input_slider(1, 10, value=3, step=0.5, label="Bin Width"))
```

### WeeksWorkedInYear: Partially-employed individuals are surprisingly rare.

Workers appear to not work at all or report 52 weeks of work.

```{r echo=F, fig.height=7}
census_sample %>% 
  ggvis(~WeeksWorkedInYear, fill=~Target) %>%
  group_by(Target) %>%
  layer_histograms(width=input_slider(1, 10, value=2, step=0.5, label="Bin Width"))
```

## Conclusion

The univariate analysis mostly matches expectations. Income is heavily dependent on an individual's education, ocupation and age. I was suprising by the lack of partiall-employed individuals. I thought there would be more workers who worked 6 months out of the year.

On to the predictive modeling!